---
title: "N allocation analysis"
author: "Jason Fridley"
date: "6/8/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TO DO:
Need defensive chemistry data to calculate Nd.
Incorporate solubles (review Takashima paper, which components for these?).
Need to do remaining calculations mentioned in proposl (see notes 6-8-22).
Put all in HB model.

## Dataset

The dataset (dat) is the merged output of the Bayesian photosynthesis model (from the script 'photo_params_all_HB4.R', including posterior means and SDs of Asat (A at 40 Pa and saturating light), Vcmax, Jmax, alpha, and Rd) and lab assay results in the master Google Docs spreadsheet (currently version 5), plus a separate csv file of associated covariates for each sample (nativity, range, growth form).

```{r}
dfold = "/Users/etudiant/Documents/IOS_git/IOS_data/"
hb = read.csv(paste0(dfold,"indoutHB4.csv"))
master = read.csv(paste0(dfold,"NSF-IOSspreadsheetJDF5.csv"))
dat = merge(hb,master,by.x="ID",by.y="ID",all=T)
dat$region = as.factor(substr((dat$Region),1,1))
covar = read.csv(paste0(dfold,"covariates.csv")) #covariates at the species-region level
dat = merge(dat,covar,by.x=c("species","region"),by.y=c("species","region"),all.x=T)
temp.dat = read.csv(paste0(dfold,"leaftempsbyID.csv")) #file of median leaf temps per sample
names(temp.dat) = c("ID","Tleaf")
dat = merge(dat,temp.dat,by.x="ID",by.y="ID",all.x=T) #merge leaf temps with dataset
```

### Colors for plotting

```{r}
#colors    
  makeTransparent<-function(someColor, alpha=150)
  {   newColor<-col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
                                              blue=curcoldata[3],alpha=alpha, maxColorValue=255)}) }
c1 = makeTransparent("darkgreen")
c2 = makeTransparent("blue")
c3 = makeTransparent("red2")
```

## Leaf N pools

We are interested in 5 pools of leaf N:

1. Nr = N in Rubisco

2. Nl = N allocated to light harvesting (Chl, corrected by 'Cb', a binding coefficient that describes the efficiency at which Nl contributes to light energy capture; Niinemets & Tenhunen 1997)

3. Nb = N allocated to bioenergetics associated with electron transfer proteins

4. Nc = N associated with cell wall proteins

5. Nd = N associated with defensive secondary metabolites, esp. cyanogenic glycosides and alkaloids

6. ?? 'Other protein nitrogen': (soluble protein N + SDS protein N) - (Nr + Nl + Nb): Takashima et al. 2004

We expect the sum of these 5 pools to be less than total leaf N, because there are residual unmeasured pools (organic acids, inorganic N). 

### N in Rubisco (Nr)

Evans (1989) reported that leaf N allocation to Rubisco varies from 5-30%, although increasing with leaf N content. To calculate Nr in our data, we note that measurements were from leaf discs and thus area based (ug cm-2), which can be transformed to mass based (g-1) by multiplying by SLA. Rubisco is 16% N, so ug N in Rubisco per g leaf is calculated by dividing Rubisco by 16. To calculate allocation (Nr = Rubisco N / leaf N), we divide by 10^6 to convert to g, and divide by g leaf N:

```{r warning=F, fig.width=8, fig.height=4}
dat$SLA = as.numeric(as.character(dat$Specific.leaf.area)) #SLA in cm2 g-1
dat$Rub.g = as.numeric((dat$Rubisco.content)) * dat$SLA #ug Rubisco per g leaf
dat$RubN = dat$Rub.g / 16 #ug N in Rubisco per g leaf
dat$Nmass = (as.numeric(as.character(dat$leafN))/100) #g leaf N per g leaf
dat$Rub.perc = (dat$RubN / 1000000) / dat$Nmass #Rubisco g per g leaf N
hist(dat$Rub.perc[dat$Rub.perc<1])
summary(dat$Rub.perc) #just a few outliers; median is 12%
par(mfrow=c(1,2),mar=c(5,5,1,1))
plot(dat$Nmass,dat$Rub.perc,col=c(c1,c2)[dat$woody+1],pch=19,ylim=c(0,.6),xlab="Nmass",ylab="N Rubisco")
legend(legend=c("Herbaceous","Woody"),bty="n",cex=.8,text.col=c("darkgreen","blue"),"topright")
plot(dat$Nmass,dat$Rub.perc,col=c(c1,c2,c3)[as.numeric(dat$Region)],pch=19,ylim=c(0,.6),xlab="Nmass",ylab="N Rubisco")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c("darkgreen","blue","red2"),"topright")

```

There are several outliers that merit inspection, but the majority of values are within the Evans range and there is not a strong bias by growth form. Nr values by region are overlapping.

### N in light harvesting (Nl)

Evans (1989) suggests that N in chl is tightly connected with total thylakoid N, such that Nl is proportional to Nb, and that together, total thylakoid N increases from about 20% in sun to over 40% in shade. His estimate is that the N cost of thylakoids is 50 mol N per mol chl, but this includes bioenergetic proteins in addition to chl itself, which scales only 4 mol N per mol chl. Thus, only 4/50 (8%) of thylakoid N is actually contained with chl, or 1.6-3.2% of total leaf N.

Niinemets & Tenhunen (1997) also consider Nl to be proportional to leaf chl content, and calculate Nl as:

Nl = chl / (Nmass * Cb)

where chl is leaf chl concentration in mmol g-1, Nmass is leaf N (g g-1), and Cb is a 'chl binding' coefficient (mmol chl per g N) that--I think--suggests some N in chloroplasts does not actually contribute to light capture; in Acer saccharum values of Cb are between 2 and 2.5. 

Takashima et al. (2004) assume that Nl is calculated based on a ratio of 37.1 mol Nl per mol chl, which is equivalent to Niinemets & Tenhunen (1997) for Nl under the assumption that Cb=1.93 (based on empirical relationship below). Note large difference in estimated Nl for maple in Niinemets & Tenhunen (1997) (20-60%) vs. oaks in Takashima (10%).

We can first examine the distribution of leaf chl content. Our measurements were based on leaf discs, and so can be converted from ug chl per cm2 to ug chl per g; assuming chl is 6.27% of chl on a mass basis (chlA = C55 H72 O5 N4 Mg, mol weight of 893.5 g/mol) we can convert to chl N. 

```{r}
dat$chl.ugcm2 = as.numeric(as.character(dat$Total.Chlorophyll)) #ug chl per cm2
dat$chl.mmolm2 = (dat$chl.ugcm2 / 1000000) * 1000 * 10000 / 893.5 #mmol chl per m2 leaf
hist(dat$chl.mmolm2) #between 0 and 0.6, consistent with Evans 1989
dat$chl.ugg = dat$chl.ugcm2 * dat$SLA  #ug chl per g leaf
dat$chlN = dat$chl.ugg / (1/.0627) #ug N in Chl per g leaf
dat$Chl.perc = (dat$chlN / 1000000) / dat$Nmass
hist(dat$Chl.perc[dat$Chl.perc<.1])
summary(dat$Chl.perc)

#examine across regions
dat$Year = as.numeric(as.character(dat$year))
par(mfrow=c(1,1),mar=c(8,4,1,1))
boxplot(chl.mmolm2~Year + Region,data=dat,las=3,xlab="")
  #problem with France values from 2020 and 2021
```

Chl content values closely match Evans 1989, and conversion to N in Chl shows a range of .01-.05 with a median of 1.4%, reasonably close to Evans' (1989) prediction. To convert to Nl, we can use either the Niinemets & Tenhunen (1997) or Takashima et al. (2004) approach, which are (nearly) equivalent if Cb=1.93 is assumed.

```{r fig.width=8, fig.height=4}
dat$chl.mmolg = dat$chl.mmolm2*dat$SLA/10000 #mmol chl per g leaf
dat$Nl.NT = dat$chl.mmolg / (dat$Nmass*1.93) #N&T '97 approach, assuming Cb is 1.93
hist(dat$Nl.NT[dat$Nl.NT<1])
  #majority between 5-30%, good correspondence to A. saccharum in N&T '97
dat$Nl.Taka = (dat$chl.mmolg * 37.1 / 1000) * 14 * (1/dat$Nmass)
hist(dat$Nl.Taka[dat$Nl.Taka<.4]) #mirrors N&T '97
summary(dat$Nl.Taka)
par(mfrow=c(1,1),mar=c(5,5,1,1))
plot(dat$Nl.Taka,dat$Nl.NT,col="gray",pch=19,xlab="Nl, Takashima",ylab="Nl, Niinemets")
abline(0,1)
```

The methods are equal, suggesting confidence in Nl estimates, and they vary widely. We expect shade plants (here, woodies) to have generally higher values of Nl. There are also outliers that need inspection.

```{r}
boxplot(Nl.Taka~woody,dat,col=c("darkgreen","blue"),ylim=c(0,.6))
```  

## N in bioenergetics (Nb)

Most authors calculate Nb in some way from Jmax. For example, Takashima et al. (2004) calculated Nb from Jmax measured from leaf samples at 25C, assuming that the ratio of Jmax to the cytochrome f content is 156 mmol mol-1 s-1 and that the N in bioenergetics per unit cytochrome f is 9.53 mol mmol-1 (see Hikosaka & Terashima 1995). Nb can thus be calculated as:

Nb (Takashima) = (Jmax/1000) * (9.53/156) * (SLA/10000) * 14 * (1/Nmass)

where Jmax is measured in umol e- m-2 s-2, SLA is cm2 g-1, and 14 converts mol N to g N. 

Niinemets & Tenhunen (1997) is similar but more general because constants include temperature dependencies of J:

Nb (N&T '97) = Jmax * (SLA/10000) * (1/Nmass) * (1/(8.06*Jmc))

where Jmc is the specific activity of electron transport in umol e- per umol cytochrome f s-1 and depends on leaf temperature; Fig. A2 in N&T '97 shows that Jmc varies between 150 and 200 from 25 to 35C and declines thereafter. The equation for temperature specificity (same as for Vmc) given by N&T is:

Jmc = exp(c-(dHa/(R*Tk))) / (1+ exp((dS*T-dHd)/R*Tk) )

The constants given for Jmc are c=14.77, dHa=24100, dHd=564150, dS=1810, R=8.314, and Tk is leaf temp. We can write a function and check for consistency with Fig. A2, and confirm that Takashima's value of 156 conforms to 25 C:

```{r}
Jmc.temp = function(c,dHa,dHd,dS,R,Tk) {
  exp(c-(dHa/(R*Tk))) / (1 + exp(  ((dS*Tk)-dHd) / (R*Tk)) )
}
Tk.vec = seq(20,40,length=100) + 273.15
y = Jmc.temp(c=14.77, dHa=24100, dHd=564150, dS=1810, R=8.314, Tk=Tk.vec)
plot(Tk.vec-273,y)
abline(v=c(25))
Jmc.temp(c=14.77, dHa=24100, dHd=564150, dS=1810, R=8.314, Tk=25+273.15)
Jmc.temp(c=14.77, dHa=24100, dHd=564150, dS=1810, R=8.314, Tk=30+273.15)
  #value at 30C should be 182
```

Results are consistent with N&T and Takashima. Note the leaf temperatures in our study were mostly taken at 30 C but vary between 25 and 33 or so, so we can vary Jmc with temp as appropriate.

```{r}
summary(dat$Tleaf) #31 NAs, assume 30C
dat$Tleaf[is.na(dat$Tleaf)] = 30
dat$Jmc = Jmc.temp(c=14.77, dHa=24100, dHd=564150, dS=1810, R=8.314, Tk=dat$Tleaf+273.15)
hist(dat$Jmc)
dat$Nb.Taka = dat$Jmax*(9.53/dat$Jmc)*(dat$SLA/10000)*14*(1/dat$Nmass)/1000
hist(dat$Nb.Taka[dat$Nb.Taka<1])
summary(dat$Nb.Taka) #average of 10%
dat$Nb.NT = dat$Jmax * (dat$SLA/10000) * (1/dat$Nmass) * (1/(8.06*dat$Jmc))
hist(dat$Nb.NT[dat$Nb.NT<1])
summary(dat$Nb.NT) #average of 10%
par(mfrow=c(1,1),mar=c(5,5,1,1))
plot(dat$Nb.Taka,dat$Nb.NT,col="gray",pch=19,xlab="Nl, Takashima",ylab="Nl, Niinemets")
abline(0,1)
  
```

Here Nb is derived solely from Jmax and assumptions about N content of bioenergetic proteins, but values centered on 10% are not unreasonable, and the methods differ only in the value of the constant and by a small amount. We can now compare Nl and Nb and confirm the expectations that 1) Nl is generally larger (Takashima et al. 2004) and 2) allocation to Nl is skewed high for shade plants (here, woodies).

```{r}
par(mfrow=c(1,1),mar=c(5,5,1,1))
plot(dat$Nl.Taka,dat$Nb.Taka,col=c(c1,c2)[dat$woody+1],pch=19,xlim=c(0,.5),ylim=c(0,.5),xlab="N allocation to light harvesting",ylab="N allocation to bioenergetics")
legend(legend=c("Herbaceous","Woody"),bty="n",cex=.8,text.col=c("darkgreen","blue"),"topleft")
abline(0,1)
```

## N in the cell wall (Nc)

We have empirical measurements of cell wall mass (in mg of the final pellet sample) and cell wall %N. From these Nc is calculated as:

Nc = [cell wall mass per leaf area] x [%N of cell wall] / [leaf Narea]

```{r}
dat$wallN = dat$Cell.wall..N / 100 #g N per g cell wall
hist(dat$wallN)
summary(dat$wallN) #mean is just under 1%, surprisingly high
dat$Narea.cm2 = dat$Nmass / dat$SLA #g N per cm2 leaf area
dat$wallg.percm2 = (dat$Cell.wall.sample.mass.mg * as.numeric(as.character(dat$cell.wall.leaf.area.cm2)))/1000 #cell wall mass, g per cm2 leaf area
dat$LMA = 1/dat$SLA #leaf mass per area, g per cm2
par(mfrow=c(1,1),mar=c(5,5,3,3))
plot(dat$LMA*1000,dat$wallg.percm2*1000,col=c(c1,c2,c3)[as.numeric(dat$Region)],pch=19,xlab="LMA mg cm-2",ylab="Cell wall mg cm-2")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c("darkgreen","blue","red2"),"bottomright")
abline(0,1); abline(0,.8,lty=2); abline(0,.6,lty=3); abline(0,.4,lty=3)
points(dat$LMA[dat$species=="Fallopia japonica"]*1000,dat$wallg.percm2[dat$species=="Fallopia japonica"]*1000,pch=21,xlab="LMA g m-2",ylab="Cell wall g m-2",col="black",lwd=2)
text(7,7,"100%",cex=.7,xpd=T)
text(9,7,"80%",cex=.7,xpd=T)
text(11.5,7,"60%",cex=.7,xpd=T)
text(12.2,4.7,"40%",cex=.7,xpd=T)

dat$Nc = (dat$wallmg.percm2 * dat$wallN) / (dat$Narea.cm2*1000) #percent leaf N present in cell wall
hist(dat$Nc[dat$Nc<1])

boxplot(dat$wallg.percm2/dat$LMA ~ dat$Region,subset=dat$species=="Fallopia japonica")

```

### N in defensive chemistry (Nd)

Waiting on data.

### Other protein N

Takashima et al. (2004) calculate this as "nitrogen in the water- and SDS soluble protein minus nitrogen in the photosynthetic apparatus", and it was the largest leaf N pool measured in that study.

```{r}
dat$Narea = dat$Nmass / (dat$SLA/10000)
dat$NUE = dat$Asat / dat$Narea
dat$Nphoto = dat$Rub.perc + dat$Nl.Taka + dat$Nb.Taka
#transform solubles data to g N per g leaf
dat$Water.protN.gg = ((as.numeric(as.character(dat$Water.soluble.Protein.mg.cm2))/1000) * dat$SLA) / 16
dat$SDS.protN.gg = ((as.numeric(as.character(dat$SDS.soluble.Protein.mg.cm2))/1000) * dat$SLA) / 16
dat$other.protN = (dat$Water.protN.gg/dat$Nmass) + (dat$SDS.protN.gg/dat$Nmass) - dat$Nphoto
hist(dat$other.protN)
  #ballpark?

plot((dat$Water.protN.gg/dat$Nmass),dat$Rub.perc,xlim=c(0,1),ylim=c(0,1))
plot((dat$SDS.protN.gg/dat$Nmass),(dat$Nl.Taka + dat$Nb.Taka),xlim=c(0,1),ylim=c(0,1),col=c(c1,c2,c3)[as.numeric(dat$Region)],pch=19)
plot((dat$SDS.protN.gg/dat$Nmass),(dat$Nl.Taka + dat$Nb.Taka),xlim=c(0,1),ylim=c(0,.7),col=c(c1,c2)[dat$woody+1],pch=19)
abline(0,1)
  #something funny going on with parts of the dataset...

```


### Comparing leaf N fractions in Fallopia across regions

```{r}
dat$Narea = dat$Nmass / (dat$SLA/10000)
dat$NUE = dat$Asat / dat$Narea
dat$Nthy = dat$Nl.Taka + dat$Nb.Taka
fall = dat[dat$sppcode=="fajap"&!is.na(dat$sppcode),]
boxplot(Nmass~Region,data=fall)
boxplot(Narea~Region,data=fall)
boxplot(Rub.perc~Region,data=fall)
boxplot(Nl.Taka~Region,data=fall)
boxplot(Nb.Taka~Region,data=fall)
boxplot(Nthy~Region,data=fall)
plot(fall$Nmass,fall$Rub.perc*fall$Nmass,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Nmass",ylab="N Rubisco")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c("darkgreen","blue","red2"),"topright")
abline(0,.1,lty=2)
text(.03,.0027,"10%")
  #Rubisco is always less than 10% of leaf N in the native range in Japan
  #in France, Rubisco % increases without an increase in leaf N
  #in ENA, Rubisco increases along with leaf N
plot(fall$Nmass,fall$Nl.Taka*fall$Nmass,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Nmass",ylab="N Light harvest")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c("darkgreen","blue","red2"),"bottomright")
abline(0,.15,lty=2)
text(.03,.004,"15%")
  #Nl is about 15% in native range in Japan
  #Nl increases to 20% in France, same Nmass
  #Nl stays 15% in ENA, but more leaf N
plot(fall$Nmass,fall$Nb.Taka*fall$Nmass,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Nmass",ylab="N Light harvest")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c("darkgreen","blue","red2"),"bottomright")
abline(0,.065,lty=2)
text(.03,.002,"6.5%")
  #in France and Japan, Nb is consistent at 8%
  #in ENA Nb decreases to 6.5%
  ##thus, Rubisco increases relative to Nb in ENA

library(RColorBrewer)
display.brewer.pal(n = 8, name = 'Greens')
bar.cols = brewer.pal(n = 8, name = 'Greens')[c(8,7,5,1)]

#create stacked bar graph of abs N in each partition, based on per g leaf
FNr.m = tapply(fall$Rub.perc*fall$Nmass,fall$Region,function(x)mean(x,na.rm=T))
FNl.m = tapply(fall$Nl.Taka*fall$Nmass,fall$Region,function(x)mean(x,na.rm=T))
FNb.m = tapply(fall$Nb.Taka*fall$Nmass,fall$Region,function(x)mean(x,na.rm=T))
Ntot.m = tapply(fall$Nmass,fall$Region,function(x)mean(x,na.rm=T))
barht = rbind(FNr.m,FNl.m,FNb.m)
tot3 = colSums(barht)
FN.miss = Ntot.m - tot3 
barht2 = rbind(barht,FN.miss)
midpoints = apply(barht2,2,cumsum) - barht2/2  #height of middle of each component for plotting
percents = (barht2 / colSums(barht2))*100 #percentages of each for plotting
perc.lab = paste0(round(percents,1),"%")
par(mar=c(3,4,1,6),mfrow=c(1,1))
b1 = barplot(barht2*1000,beside=F,ylab="mg N per N leaf",col=bar.cols)
text(rep(b1,each=4),as.numeric(midpoints)*1000,perc.lab,col="white",cex=.8,lwd=2)
mtext("Rubisco",side=4,las=1,at=midpoints[1,3]*1000)
mtext("Light harvest",side=4,las=1,at=midpoints[2,3]*1000)
mtext("Bioenergetics",side=4,las=1,at=midpoints[3,3]*1000)

#create stacked bar graph of abs N in each partition, based on per m2 leaf
FNr.m = tapply(fall$Rub.perc*fall$Narea,fall$Region,function(x)mean(x,na.rm=T))
FNl.m = tapply(fall$Nl.Taka*fall$Narea,fall$Region,function(x)mean(x,na.rm=T))
FNb.m = tapply(fall$Nb.Taka*fall$Narea,fall$Region,function(x)mean(x,na.rm=T))
Ntot.m = tapply(fall$Narea,fall$Region,function(x)mean(x,na.rm=T))
barht = rbind(FNr.m,FNl.m,FNb.m)
tot3 = colSums(barht)
FN.miss = Ntot.m - tot3 
barht2 = rbind(barht,FN.miss)
midpoints = apply(barht2,2,cumsum) - barht2/2  #height of middle of each component for plotting
percents = (barht2 / colSums(barht2))*100 #percentages of each for plotting
perc.lab = paste0(round(percents,1),"%")
par(mar=c(3,4,1,6))
b1 = barplot(barht2,beside=F,ylab="g N per m2 leaf",col=bar.cols)
text(rep(b1,each=4),as.numeric(midpoints),perc.lab,col="white",cex=.8,lwd=2)
mtext("Rubisco",side=4,las=1,at=midpoints[1,3])
mtext("Light harvest",side=4,las=1,at=midpoints[2,3])
mtext("Bioenergetics",side=4,las=1,at=midpoints[3,3])


#Asat vs. Rubisco vs. Narea 
par(mfrow=c(2,2),mar=c(4,4,1,1))
plot(fall$Narea,fall$Asat,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Leaf N (g m-2)",ylab="Asat")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c("darkgreen","blue","red2"),"bottomright")
plot(fall$Rub.perc*fall$Narea,fall$Asat,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Rubisco N (g m-2)",ylab="Asat")
plot(fall$Nl.Taka*fall$Narea,fall$Asat,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Light harvest N (g m-2)",ylab="Asat")
plot(fall$Nb.Taka*fall$Narea,fall$Asat,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Bioenergetics N (g m-2)",ylab="Asat")
  #this is because Jmax drives Asat in this species at 40 Pa Ci and saturating light
  #plot(fall$Vcmax,fall$Asat)
summary.aov(lm(Asat~Narea*Region,data=fall))
TukeyHSD(aov(Asat~Narea*Region,data=fall),which="Region") #only weak evidence of difference

#Rubisco should better correlate with Vcmax
par(mfrow=c(1,2),mar=c(4,4,1,1))
plot(fall$Narea,fall$Vcmax,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Leaf N (g m-2)",ylab="Vcmax")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c("darkgreen","blue","red2"),"topleft")
plot(fall$Rub.perc*fall$Narea,fall$Vcmax,col=c(c1,c2,c3)[as.numeric(fall$Region)],pch=19,xlab="Rubisco N (g m-2)",ylab="Vcmax")
  #but it doesn't; suggests a problem with Rubisco?

#PNUE
boxplot(NUE~Region,data=fall)
summary.aov(lm(NUE~Region,data=fall))
TukeyHSD(aov(NUE~Region,data=fall)) #ENA lower PNUE than both Japan or France

#plot Nl vs. Nthy across 3 groups


```

### alpha vs. chl

#from N&T '97
```{r}
dat$LMA = 1/(dat$SLA/10000) #dry mass per unit area in g m-2
alpha.pred = .292 * ((1 +(.076/(dat$chl.mmolg*dat$LMA)))^(-1))
plot(alpha.pred,dat$alpha,ylim=c(0,1)); abline(0,1)
hist(dat$alpha[dat$alpha<1])
summary(lm(dat$alpha~alpha.pred,subset=dat$alpha<1))
  #quite a terrible relationship!
```

