---
title: "N allocation analysis"
author: "Jason Fridley"
date: "6/8/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset

The dataset (dat) is the merged output of the Bayesian photosynthesis model (from the script 'photo_params_all_HB4.R', including posterior means and SDs of Asat (A at 40 Pa and saturating light), Vcmax, Jmax, alpha, and Rd) and lab assay results in the master Google Docs spreadsheet (currently version 5), plus a separate csv file of associated covariates for each sample (nativity, range, growth form).

```{r}
dfold = "/Users/etudiant/Documents/IOS_git/IOS_data/"
hb = read.csv(paste0(dfold,"indoutHB4.csv"))
master = read.csv(paste0(dfold,"NSF-IOSspreadsheetJDF5.csv"))
dat = merge(hb,master,by.x="ID",by.y="ID",all=T)
dat$region = as.factor(substr((dat$Region),1,1))
covar = read.csv(paste0(dfold,"covariates.csv")) #covariates at the species-region level
dat = merge(dat,covar,by.x=c("species","region"),by.y=c("species","region"),all.x=T)
```

## Colors for plotting

```{r}
#colors    
  makeTransparent<-function(someColor, alpha=150)
  {   newColor<-col2rgb(someColor)
  apply(newColor, 2, function(curcoldata){rgb(red=curcoldata[1], green=curcoldata[2],
                                              blue=curcoldata[3],alpha=alpha, maxColorValue=255)}) }
c1 = makeTransparent("darkgreen")
c2 = makeTransparent("blue")
c3 = makeTransparent("red2")
```

## Leaf N pools

We are interested in 5 pools of leaf N:

1. Nr = N in Rubisco

2. Nl = N allocated to light harvesting (Chl, corrected by 'Cb', a binding coefficient that describes the efficiency at which Nl contributes to light energy capture; Niinemets & Tenhunen 1997)

3. Nb = N allocated to bioenergetics associated with electron transfer proteins

4. Nc = N associated with cell wall proteins

5. Nd = N associated with defensive secondary metabolites, esp. cyanogenic glycosides and alkaloids

We expect the sum of these 5 pools to be less than total leaf N, because there are residual unmeasured pools (organic acids, inorganic N). 

### N in Rubisco (Nr)

Evans (1989) reported that leaf N allocation to Rubisco varies from 5-30%, although increasing with leaf N content. To calculate Nr in our data, we note that measurements were from leaf discs and thus area based (ug cm-2), which can be transformed to mass based (g-1) by multiplying by SLA. Rubisco is 16% N, so ug N in Rubisco per g leaf is calculated by dividing Rubisco by 16. To calculate allocation (Nr = Rubisco N / leaf N), we divide by 10^6 to convert to g, and divide by g leaf N:

```{r warning=F, fig.width=8, fig.height=4}
dat$SLA = as.numeric(as.character(dat$Specific.leaf.area)) #SLA in cm2 g-1
dat$Rub.g = as.numeric((dat$Rubisco.content)) * dat$SLA #ug Rubisco per g leaf
dat$RubN = dat$Rub.g / 16 #ug N in Rubisco per g leaf
dat$Nmass = (as.numeric(as.character(dat$leafN))/100) #g leaf N per g leaf
dat$Rub.perc = (dat$RubN / 1000000) / dat$Nmass #Rubisco g per g leaf N
hist(dat$Rub.perc[dat$Rub.perc<1])
summary(dat$Rub.perc) #just a few outliers; median is 12%
par(mfrow=c(1,2),mar=c(5,5,1,1))
plot(dat$Nmass,dat$Rub.perc,col=c(c1,c2)[dat$woody+1],pch=19,ylim=c(0,.6),xlab="Nmass",ylab="N Rubisco")
legend(legend=c("Herbaceous","Woody"),bty="n",cex=.8,text.col=c("darkgreen","blue"),"topright")
plot(dat$Nmass,dat$Rub.perc,col=c(c1,c2,c3)[as.numeric(dat$Region)],pch=19,ylim=c(0,.6),xlab="Nmass",ylab="N Rubisco")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c("darkgreen","blue","red2"),"topright")

```

There are several outliers that merit inspection, but the majority of values are within the Evans range and there is not a strong bias by growth form. Nr values by region are overlapping.

### N in light harvesting (Nl)

Evans (1989) suggests that N in chl is tightly connected with total thylakoid N, such that Nl is proportional to Nb, and that together, total thylakoid N increases from about 20% in sun to over 40% in shade. His estimate is that the N cost of thylakoids is 50 mol N per mol chl, but this includes bioenergetic proteins in addition to chl itself, which scales only 4 mol N per mol chl. Thus, only 4/50 (8%) of thylakoid N is actually contained with chl, or 1.6-3.2% of total leaf N.

Niinemets & Tenhunen (1997) also consider Nl to be proportional to leaf chl content, and calculate Nl as:

Nl = chl / (Nmass * Cb)

where chl is leaf chl concentration in mmol g-1, Nmass is leaf N (g g-1), and Cb is a 'chl binding' coefficient (mmol chl per g N) that--I think--suggests some N in chloroplasts does not actually contribute to light capture; in Acer saccharum values of Cb are between 2 and 2.5. 

Takashima et al. (2004) assume that Nl is calculated based on a ratio of 37.1 mol Nl per mol chl. Note large difference in estimated Nl for maple in Niinemets & Tenhunen (1997) (20-60%) vs. oaks in Takashima (10%).

We can first examine the distribution of leaf chl content. Our measurements were based on leaf discs, and so can be converted from ug chl per cm2 to ug chl per g; assuming chl is 6.27% of chl on a mass basis (chlA = C55 H72 O5 N4 Mg, mol weight of 893.5 g/mol) we can convert to chl N. 

```{r}
dat$chl.ugcm2 = as.numeric(as.character(dat$Total.Chlorophyll)) #ug chl per cm2
dat$chl.mmolm2 = (dat$chl.ugcm2 / 1000000) * 1000 * 10000 / 893.5 #mmol chl per m2 leaf
hist(dat$chl.mmolm2) #between 0 and 0.6, consistent with Evans 1989
dat$chl.ugg = dat$chl.ugcm2 * dat$SLA  #ug chl per g leaf
dat$chlN = dat$chl.ugg / (1/.0627) #ug N in Chl per g leaf
dat$Chl.perc = (dat$chlN / 1000000) / dat$Nmass
hist(dat$Chl.perc[dat$Chl.perc<.1])
summary(dat$Chl.perc)
```

Chl content values closely match Evans 1989, and conversion to N in Chl shows a range of .01-.05 with a median of 1.4%, reasonably close to Evans' (1989) prediction. To convert to Nl, we can use either the Niinemets & Tenhunen (1997) or Takashima et al. (2004) approach:

```{r fig.width=8, fig.height=4}
dat$chl.mmolg = dat$chl.mmolm2*dat$SLA/10000 #mmol chl per g leaf
dat$Nl.NT = dat$chl.mmolg / (dat$Nmass*2.3) #N&T '97 approach, assuming Cb is 2.3
hist(dat$Nl.NT[dat$Nl.NT<1])
  #majority between 5-30%, good correspondence to A. saccharum in N&T '97
dat$Nl.Taka = (dat$chl.mmolg * 37.1 / 1000) * 14 #Takashima et al. 2004 approach
hist(dat$Nl.Taka[dat$Nl.Taka<.4]*10) #most between 2-5%, median is 3%, somewhat lower than oaks in Takashima (~10%)
summary(dat$Nl.Taka)
par(mfrow=c(1,2),mar=c(5,5,1,1))
plot(dat$Nl.Taka,dat$Nl.NT,col=c(c1,c2)[dat$woody+1],pch=19,ylim=c(0,.45),xlim=c(0,.2),xlab="Nl, Takashima",ylab="Nl, Niinemets")
legend(legend=c("Herbaceous","Woody"),bty="n",cex=.8,text.col=c("darkgreen","blue"),"bottomright")
abline(0,1)
plot(dat$Nl.Taka,dat$Nl.NT,col=c(c1,c2,c3)[as.numeric(dat$Region)],pch=19,ylim=c(0,.45),xlim=c(0,.2),xlab="Nl, Takashima",ylab="Nl, Niinemets")
legend(legend=c("ENA","France","Japan"),bty="n",cex=.8,text.col=c("darkgreen","blue","red2"),"bottomright")
abline(0,1)

```


